---
title: Radically Infinite
author: Derek Sollberger
date: '2022-07-06'
banner: "img/radically_infinte.png"
slug: radically-infinite
categories:
  - math
tags:
  - latex
---

In today's procrastination effort, I want to generate the `LaTeX` code to produce an infinite radical expansion for the number 3, and generate the code using `R`.

... inspired by math teacher Chris Smith

* [tweet](https://mobile.twitter.com/aap03102/status/1544265133735690242)

First, let me see if I can simply generate the pattern of integers.  I will use the letter capital "S" and later replace it with the square root symbol.

```{r}
stem <- "S 1 + "
print(paste(stem, "8"))
for(a in 2:5){
  b <- (a+1)^2-1
  print(paste(stem, a, "*", b/a))
  print(paste(stem, a, "S", (b/a)^2))
  print(paste(stem, a, "S", 1, "+", (b/a)^2- 1))
  stem <- paste(stem, a, "S", 1, "+")
}
```

Not only do I need to now insert the square root symbol, but I would also need to close the parentheses.  I think I an literally count the number of times "S" appears and work with that.

```{r}
stem <- "S1 + "
print("\\sqrt(9)")
print("\\sqrt(1 + 8)")
for(a in 2:5){
  b <- (a+1)^2-1
  
  line1 <- paste(stem, a, "*", b/a)
  n1 <- stringr::str_count(line1, "S")
  line1 <- stringr::str_replace_all(line1, "S", "\\\\sqrt(")
  line1 <- paste(line1, paste0(rep(")", n1), collapse = ""), "\\\\")
  print(line1)
  
  line2 <- paste(stem, a, "S", (b/a)^2)
  n2 <- stringr::str_count(line2, "S")
  line2 <- stringr::str_replace_all(line2, "S", "\\\\sqrt(")
  line2 <- paste(line2, paste0(rep(")", n2), collapse = ""), "\\\\")
  print(line2)
  
  line3 <- paste(stem, a, "S", 1, "+", (b/a)^2- 1)
  n3 <- stringr::str_count(line3, "S")
  line3 <- stringr::str_replace_all(line3, "S", "\\\\sqrt(")
  line3 <- paste(line3, paste0(rep(")", n3), collapse = ""), "\\\\")
  print(line3)
  
  stem <- paste0(stem, a, "S", 1, "+")
}
```

Now, I want to move toward a LaTeX `array` environment.

```{r}
print("\\begin{array}{rcl}")
stem <- "S1 + "
print("3 & = & \\sqrt(9)")
print("3 & = & \\sqrt(1 + 8)")
for(a in 2:5){
  b <- (a+1)^2-1
  
  line1 <- paste(stem, a, "*", b/a)
  n1 <- stringr::str_count(line1, "S")
  line1 <- stringr::str_replace_all(line1, "S", "\\\\sqrt(")
  line1 <- paste(line1, paste0(rep(")", n1), collapse = ""), "\\\\")
  print(paste("3 & = &", line1))
  
  line2 <- paste(stem, a, "S", (b/a)^2)
  n2 <- stringr::str_count(line2, "S")
  line2 <- stringr::str_replace_all(line2, "S", "\\\\sqrt(")
  line2 <- paste(line2, paste0(rep(")", n2), collapse = ""), "\\\\")
  print(paste("3 & = &", line2))
  
  line3 <- paste(stem, a, "S", 1, "+", (b/a)^2- 1)
  n3 <- stringr::str_count(line3, "S")
  line3 <- stringr::str_replace_all(line3, "S", "\\\\sqrt(")
  line3 <- paste(line3, paste0(rep(")", n3), collapse = ""), "\\\\")
  print(paste("3 & = &", line3))
  
  stem <- paste0(stem, a, "S", 1, "+")
}
print("\\end{array}")
```

Now let's see if we can collect everything into just one string (of LaTeX code)

```{r}
main_string <- "\\$\\begin{array}{rcl}"
stem <- "S1 + "
main_string <- paste(main_string, "3 & = & \\sqrt{9} \\\\")
main_string <- paste(main_string, "3 & = & \\sqrt{1 + 8} \\\\")
for(a in 2:7){
  b <- (a+1)^2-1
  
  line1 <- paste(stem, a, "*", b/a)
  n1 <- stringr::str_count(line1, "S")
  line1 <- stringr::str_replace_all(line1, "S", "\\\\sqrt{")
  line1 <- paste(line1, paste0(rep("}", n1), collapse = ""), "\\\\")
  main_string <- paste(main_string, "3 & = &", line1)
  
  line2 <- paste(stem, a, "S", (b/a)^2)
  n2 <- stringr::str_count(line2, "S")
  line2 <- stringr::str_replace_all(line2, "S", "\\\\sqrt{")
  line2 <- paste(line2, paste0(rep("}", n2), collapse = ""), "\\\\")
  main_string <- paste(main_string, "3 & = &", line2)
  
  line3 <- paste(stem, a, "S", 1, "+", (b/a)^2- 1)
  n3 <- stringr::str_count(line3, "S")
  line3 <- stringr::str_replace_all(line3, "S", "\\\\sqrt{")
  line3 <- paste(line3, paste0(rep("}", n3), collapse = ""), "\\\\")
  main_string <- paste(main_string, "3 & = &", line3)
  
  stem <- paste0(stem, a, "S", 1, "+")
}
main_string <- paste(main_string, "\\end{array}\\$")
```

Finally, does this print in inline LaTeX handling here in RStudio??

`r main_string`

Oh well.  At least I did generate replicable code that can be ported over to a LaTeX editor!