---
title: Making Flash Cards
author: Derek Sollberger
date: '2022-07-07'
banner: "img/Day11b.jpg"
slug: making-flash-cards
categories:
  - publishing
tags: []
---

Today I wanted to try out the `flashr` package for making flash cards (i.e. see if this is something I want to pursue further).

Someone (I am assuming his name is Jeffrey Stevens) over in the R4DS Slack channel mentioned a [blog post](https://jeffreyrstevens.github.io/flashr/) using the `flashr` package.


```{r}
# https://jeffreyrstevens.github.io/flashr/
#devtools::install_github("JeffreyRStevens/flashr")
library("dplyr")
library("flashr")
```

From one of my `30 Day Map Challenge` adventures, I had a data frame about availability of therapy in the United States.  

```{r}
df <- readxl::read_xlsx("counseling_ranks.xlsx")
```

Today, I am going to make myself a quiz on states and their abbreviations.

```{r}
df_info <- df |>
  select(region, id) |>
  
  #to fit the flashr mechanisms
  rename(term = region, description = id) |>
  mutate(package = "")
```

```{r}
readr::write_csv(df_info, "my_deck.csv")
```

Finally, let's see if this works!

```{r}
flashr::flashcard("my_deck.csv")
```

Neat!  Hopefully, the `flashr` package has more features in the future.
