---
title: Pascal's Triangle
author: Derek Sollberger
date: '2022-07-08'
banner: "pascal_plot.gif"
slug: pascal-s-triangle
categories:
  - data visualization
tags:
  - ggplot
---

Once again, I find that I have been nerd sniped (inspired by [Mathigon](https://mobile.twitter.com/MathigonOrg/status/1545407481119211520)).  Today, I want to plot (and make an animation) for Pascal's Triangle mod `n`.  What I am thinking is

* store Pascal Triangle numbers along with row and column numbers in a data frame
* then assign `x` and `y` values
* *rescale* the `x` and `y` values to make an equilateral triangle

```{r, message = FALSE, warning = FALSE}
# load package(s)
library("tidyverse")
```


```{r}
# allocate data space
N <- 50 #number of rows of wanted Pascal's Triangle
# EDIT: my computer or software did not seem to handle larger numbers well
row_total <- N*(N+1) / 2 #sum of 1 + 2 + ... + N

row_num <- rep(0, row_total)
col_num <- rep(0, row_total)
PT_val  <- rep(0, row_total)
x       <- rep(0, row_total)
y       <- rep(0, row_total)

df <- data.frame(row_num, col_num, PT_val, x, y)
```

Let us pretend we are back in our high school programming class and are asked to code up Pascal's Triangle.  [What's missing here is all of my simple errors.]

```{r}
# compute row and column numbers for data frame
m <- 1 #row counter in data frame
for(i in 1:N){
  for(j in 1:i){
    df$row_num[m] <- i
    df$col_num[m] <- j
    m <- m + 1
  }
}

# compute Pascal Triangle values
df$PT_val[df$row_num == 1 & df$col_num == 1] <- 1
df$PT_val[df$row_num == 2 & df$col_num == 1] <- 1
df$PT_val[df$row_num == 2 & df$col_num == 2] <- 1
for(i in 3:N){
  for(j in 1:i){
    if(j == 1){
      df$PT_val[df$row_num == i & df$col_num == j] <- 1
    } else if(j == i) {
      df$PT_val[df$row_num == i & df$col_num == j] <- 1
    } else {  
      df$PT_val[df$row_num == i & df$col_num == j] <- 
        df$PT_val[df$row_num == (i-1) & df$col_num == (j-1)] +
        df$PT_val[df$row_num == (i-1) & df$col_num == j]
    }
  }
}
```

The $y$ values are easy.  I simply need to reverse the order of the row numbers.

```{r}
df$y <- N - df$row_num + 1
```

There is probably a clever way of creating $x$ values (toward a visual plot), but I am thinking of a brute force process

* row $m$ has $m$ non-zero Pascal's Triangle numbers in it
* we can make a simple array 1, 2, ..., $m$ [actually, going to count by twos]
* then shift the array so that it is *centered* at $x = 0$

```{r}
for(i in 1:N){
  these_x_vals <- seq(1, 2*i, 2) - i
  for(j in 1:i){
    df$x[df$row_num == i & df$col_num == j] <- these_x_vals[j]
  }  
}
```

I want my graphs to make an equilateral triangle, so I need to rescale the $x$ values by a factor of

$$\frac{1}{2} \div \frac{\sqrt{3}}{2} = \frac{1}{2} * \frac{2}{\sqrt{3}} = \frac{1}{\sqrt{3}}$$

```{r}
df$x <- df$x / sqrt(3)
```

Let's make a graph!

```{r}
df |>
  ggplot() +
  geom_point(aes(x = x, y = y)) +
  coord_equal() #ensuring equal scales for x and y
```

Now for the nerdy part!  We can compute the Pascal Triangle numbers modulo $n$ for some natural number $n$, and the truly divisible numbers create fractal patterns.

```{r}
n <- 2
df |>
  ggplot() +
  geom_point(aes(x = x, y = y,
                 color = (PT_val %% n) == 0),
             size = 5) +
  coord_equal() +
  labs(title = "Pascal's Triangle",
       subtitle = paste("divisible by", n),
       caption = "Derek Sollberger") +
  scale_color_manual(values = c("gray90", "#D60270")) +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        plot.title = element_text(face = "bold", size = 20, 
                                  color = "#0038A8", hjust = 0.5),
        plot.subtitle = element_text(size = 15, 
                                     color = "#D60270", hjust = 0.5),
        plot.caption = element_text(size = 10, 
                                    color = "#0038A8", hjust = 0.5))
  
```

To make an animation, I tend to elect to make several images, and then use a 3rd-party app to combine the images.

```{r}
#| eval = FALSE
for(n in 2:20){
  this_plot <- df |>
    ggplot() +
    geom_point(aes(x = x, y = y,
                   color = (PT_val %% n) == 0),
               size = 1) +
    coord_equal() +
    labs(title = "Pascal's Triangle",
         subtitle = paste("divisible by", n),
         caption = "Derek Sollberger") +
    scale_color_manual(values = c("gray75", "#D60270")) +
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          legend.position = "none",
          panel.background = element_blank(),
          plot.title = element_text(face = "bold", size = 15, 
                                    color = "#0038A8", hjust = 0.5),
          plot.subtitle = element_text(size = 15, 
                                       color = "#D60270", hjust = 0.5),
          plot.caption = element_text(size = 10, 
                                      color = "#0038A8", hjust = 0.5))
  
  # save this plot
  ggsave(paste0("pascal_plot_", n, ".png"), plot = this_plot, 
         device = "png", height = 800, width = 1200, 
         dpi = 300, units = "px")
}
```




