---
title: Kruskal-Wallis Test
author: Derek Sollberger
date: '2022-08-10'
slug: kruskal-wallis-test
categories:
  - inference
tags:
  - ggplot
---

I want to mentally prepare for the upcoming semester.  Today, let me do a warm-up exercise of computing a `Kruskal-Wallis test`.

```{r, message = FALSE, warning = FALSE}
library("palmerpenguins")
library("tidyverse")
```

You can see that I will use the ubiquitous `Palmer Penguins` data set

```{r}
str(penguins)
```

Our null hypothesis will be that the three species of penguins have the *same* bill length.  Let us first get a visual.

```{r}
penguins |>
  ggplot() +
  geom_boxplot(aes(x = species, y = bill_length_mm,
                   fill = species)) +
  labs(title = "Palmer Penguins",
       subtitle = "null hypothesis: the three species of penguins have the *same* bill length",
       caption = "Derek Sollberger") +
  theme_minimal()
```

Visually, it appears that the penguins tend to have different bill lengths.  To be more scientifically sound, let us now run the `Kruskal-Wallis` test.

```{r}
# https://datasciencetut.com/how-to-perform-the-kruskal-wallis-test-in-r/
kruskal.test(bill_length_mm ~ species, data = penguins)
```

Since the p-value is less than 0.05, let us reject the null hypothesis of equal bill lengths.

===

The blog post mentioned above also told of a way to do pairwise comparisons in this setting (i.e. more than 2 labels)

```{r}
pairwise.wilcox.test(penguins$bill_length_mm, penguins$species,
                     p.adjust.method = "BH")
```
In this quick example, it appears that, pairwise, each species bill length is significally different than the other species.
