---
title: Trying out the Flow package
author: Derek Sollberger
date: '2022-07-06'
banner: 'img/20220706.png'
slug: trying-out-the-flow-package
categories:
  - R Markdown
tags:
  - ggplot
---

I have been meaning to try out the `flow` package as it appears to help people visualize loops and conditional statements.

Here is some code that I made to help someone over in the R4DS Slack channel. The task was to check if the change in temperature was over 0.5, and then to set the next 30 days of data to `NA`.

```{r}
temp_data <- rnorm(100, 20, 1) #fake data about temperatures
N <- length(temp_data)

#initialization
buffer <- 0
day <- 2
previous_temp <- temp_data[1]

# loop
while(day <= N){
  current_temp <- temp_data[day]
  
  if(buffer == 0){
    if(!is.na(current_temp) & !is.na(previous_temp)){
      #perform check
      if(abs(current_temp - previous_temp) > 0.5){
      buffer <- 30
    }}
  } else {
    temp_data[day] <- NA
    buffer <- buffer - 1
  }
  
  # iterate
  previous_temp <- current_temp
  day <- day + 1
}
```

For the `flow` package capabilities, I think I need to wrap this code chunk into a function.

```{r}
prune_data <- function(temp_data){
  N <- length(temp_data)
  
  #initialization
  buffer <- 0
  day <- 2
  previous_temp <- temp_data[1]
  
  # loop
  while(day <= N){
    current_temp <- temp_data[day]
    
    if(buffer == 0){
      if(!is.na(current_temp) & !is.na(previous_temp)){
        #perform check
        if(abs(current_temp - previous_temp) > 0.5){
          buffer <- 30
        }}
    } else {
      temp_data[day] <- NA
      buffer <- buffer - 1
    }
    
    # iterate
    previous_temp <- current_temp
    day <- day + 1
  }
}
```

Now let us see if the `flow` package can visualize the `prune_data` function.

```{r}
flow::flow_view(prune_data)
```
